---
title: Universe of iSEE panels
author:
- name: Aaron Lun
  email: infinite.monkeys.with.keyboards@gmail.com
date: "`r BiocStyle::doc_date()`"
package: iSEE
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Panel universe}
  %\VignetteEncoding{UTF-8}  
  %\VignettePackage{iSEEu}
  %\VignetteKeywords{GeneExpression, RNASeq, Sequencing, Visualization, QualityControl, GUI}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, echo=FALSE}
knitr::opts_chunk$set(error=FALSE, warning=FALSE, message=FALSE)
library(BiocStyle)
```

# Overview

The `r Biocpkg("iSEE")` package provides a general and flexible framework for interactively exploring `SummarizedExperiment` objects.
However, in many cases, more specialized panels are required for effective visualization of specific data types.
The `r Biocpkg("iSEEu")` package implements a collection of such dedicated panel classes that work directly in the `iSEE` application and can smoothly interact with other panels.
This allows users to quickly parametrize bespoke apps for their data to address scientific questions of interest.
We first load in the package:

```{r}
library(iSEEu)
```

All the panels described in this document can be specified by simply passing them into the `iSEE()` function via the `initial=` argument.
For example, we can create an app to explore differential expression results using a volcano and MA plot:

```{r}
# Making up some results:
se <- SummarizedExperiment(matrix(rnorm(10000), 1000, 10))
rownames(se) <- paste0("GENE_", seq_len(nrow(se)))
rowData(se)$PValue <- runif(nrow(se))
rowData(se)$LogFC <- rnorm(nrow(se))
rowData(se)$AveExpr <- rnorm(nrow(se))

app <- iSEE(se, initial=list(VolcanoPlot(), MAPlot()))
```

# Differential expression plots

The `MAPlot` class creates a MA plot, i.e., with the log-fold change on the y-axis and the average expression on the x-axis.
Features with significant differences in each direction are highlighted and counted on the legend.
Users can vary the significance threshold and apply _ad hoc_ filters on the log-fold change.
This is a subclass of the `RowDataPlot` so points can be transmitted to other panels as multiple row selections.
Instances of this class are created like:

```{r}
ma.panel <- MAPlot()
class(ma.panel)
```

The `VolcanoPlot` class creates a volcano plot with the log-fold change on the x-axis and the negative log-p-value on the y-axis.
Features with significant differences in each direction are highlighted and counted on the legend.
Users can vary the significance threshold and apply _ad hoc_ filters on the log-fold change.
This is a subclass of the `RowDataPlot` so points can be transmitted to other panels as multiple row selections.
Instances of this class are created like:

```{r}
vol.panel <- VolcanoPlot()
class(vol.panel)
```

# Session information

```{r}
sessionInfo()
```

